Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-28 Documents/Paper/ms.tex
Starfield is basically the same
Mention mixture models
Mention mixtures in abstract

\begin{document}
\maketitle
\abstract{Many inference problems involve inferring the number $N$ of
objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
from a dataset $\mathcal{D}$. These problems are typically solved using one
of the following two methods: i)
by executing a Monte Carlo algorithm (such as Nested Sampling) once for each
possible value of $N$,
and calculating the marginal likelihood or evidence as a function of $N$; or
ii) by doing a single run that allows the model dimension $N$ to change (such as
Markov Chain Monte Carlo with birth/death moves), and obtaining the posterior
for $N$ directly. In this paper we present a general approach to this problem
that uses trans-dimensional MCMC embedded {\it within} a Nested Sampling
algorithm, allowing us to explore the posterior distribution and calculate
the marginal likelihood (summed over $N$) even if the problem contains a phase
transition or other difficult features such as multimodality.
We present two example problems, finding sinusoidal signals in
noisy data, and finding and measuring galaxies in a noisy astronomical image.}

\section{Introduction}\label{sec:introduction}
\setlength{\parindent}{0cm}
\setlength{\parskip}{3mm}
Consider the following class of inference problems. There is some unknown
number, $N$, of objects in a (physical or metaphorical) region. Each of the $N$ objects has a property
$\xx$, which may be a single scalar value (for example, a mass), or a set of
values (e.g. a two-dimensional position and a mass). These problems are often
challenging to solve due to the unknown (and potentially large) dimensionality.
In addition, they also have the so-called ``label-switching degeneracy'' issue
\citep{label_switching}, where the meaning of the model is invariant to
switching the identities of the objects.

@@

%\section{``Transit'' Example}
%\section{``Asteroseismology'' Example}
\section{``Galaxy Field'' Example}
Source detection is an important problem in astrophysics. Given some data,
usually one or more noisy, blurred, images of the sky, we would like to know
how many objects $N$ (such as stars or galaxies) are in the image, and their
properties (such as flux, size, orientation). Various approaches exist,
covering a wide spectrum between ad-hoc approaches and principled inference
approaches
\citep[e.g.][]{irwin, sextractor, dolphot, 2003MNRAS.338..765H, starfield}.
The more principled techniques tend to be much more computationally
intensive, so should only be applied to interesting or
especially challenging subsets of astronomical imaging.

>>>>

\begin{document}
\maketitle
\abstract{Many inference problems involve inferring the number $N$ of
objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
from a dataset $\mathcal{D}$. A common statistical example is finite mixture
modelling. In the Bayesian framework, these problems are typically solved using
one of the following two methods: i)
by executing a Monte Carlo algorithm (such as Nested Sampling) once for each
possible value of $N$,
and calculating the marginal likelihood or evidence as a function of $N$; or
ii) by doing a single run that allows the model dimension $N$ to change (such as
Markov Chain Monte Carlo with birth/death moves), and obtaining the posterior
for $N$ directly. In this paper we present a general approach to this problem
that uses trans-dimensional MCMC embedded {\it within} a Nested Sampling
algorithm, allowing us to explore the posterior distribution and calculate
the marginal likelihood (summed over $N$) even if the problem contains a phase
transition or other difficult features such as multimodality.
We present two example problems, finding sinusoidal signals in
noisy data, and finding and measuring galaxies in a noisy astronomical image.}

\section{Introduction}\label{sec:introduction}
\setlength{\parindent}{0cm}
\setlength{\parskip}{3mm}
Consider the following class of inference problems. There is some unknown
number, $N$, of objects in a (physical or metaphorical) region. Each of the $N$ objects has a property
$\xx$, which may be a single scalar value (for example, a mass), or a set of
values (e.g. a two-dimensional position and a mass). These problems are often
challenging to solve due to the unknown (and potentially large) dimensionality.
In addition, they also have the so-called ``label-switching degeneracy'' issue
\citep{label_switching}, where the meaning of the model is invariant to
switching the identities of the objects. Finite mixture modelling is a common
statistical example.

@@

%\section{``Transit'' Example}
%\section{``Asteroseismology'' Example}
\section{``Galaxy Field'' Example}
Source detection is an important problem in astrophysics. Given some data,
usually one or more noisy, blurred, images of the sky, we would like to know
how many objects $N$ (such as stars or galaxies) are in the image, and their
properties (such as flux, size, orientation). Various approaches exist,
covering a wide spectrum between ad-hoc approaches and principled inference
approaches
\citep[e.g.][]{irwin, sextractor, dolphot, 2003MNRAS.338..765H, starfield}.
The more principled techniques tend to be much more computationally
intensive, so should only be applied to interesting or
especially challenging subsets of astronomical imaging. In fact, the approach
used by \citet{starfield} was essentially an early version of the methodology
presented in this paper.
